Manuscript progress report

### 2024-01-23 Docs/manuscript.Rmd
GFOSR simulation output

load(here("Data/SimOutput_fGFPCA.RData"))
load(here("Data/SimOutput_GLMMadaptive.RData"))

## prediction window 
window <- seq(0, 1, by = 0.2)
M <- length(pred_list_all)
```

@@

```{r, eval=FALSE}
grid_exp <- data.frame(t = seq(0, 1, by = 0.01), y = 0)
ggplot(grid_exp, aes(x=t, y=y))+
  geom_point(size = 0.5)
```

@@

@@

```{r fig_sim_data}
rand_id <- sample(501:600, size = 4)
pred_list_all[[1]] %>% filter(id %in% rand_id) %>%
  mutate_at(vars(eta_i), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"), show.legend = F)+
  facet_wrap(~id)+
  labs(title = "Simulated data")+
  scale_x_continuous(breaks = seq(0, 1, by = 0.2))
```

## Reference method

- GLMMadaptive
- Here we can fit a model with random intercept and slope for time. It is doable on 500 datasets, but obviously too simple for the data generation scheme. We would expect it to perform terribly.

@@

```{r}
# range(pred_list_all[[1]]$pred0.2, na.rm = T)
fig_fgfpca <- pred_list_all[[1]] %>%
  filter(id %in% rand_id) %>%
  mutate_at(vars(eta_i, pred0.2, pred0.4, pred0.6, pred0.8), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.6, col = "0.6"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), linetype="dashed", na.rm=T)+
  facet_wrap(~id)+
  labs(title = "fGFPCA", col = "Maximum observation time")+
  scale_x_continuous(breaks = seq(0, 1, by = 0.2))+
  theme(axis.text = element_text(size = 5))

fig_adglmm <- pred_list_ref[[1]] %>%
  filter(id %in% rand_id) %>%
  mutate_at(vars(eta_i, pred0.2, pred0.4, pred0.6, pred0.8), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.6, col = "0.6"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), linetype="dashed", na.rm=T)+
  facet_wrap(~id)+
  labs(title = "GLMMadaptive", col = "Maximum observation time")+
  scale_x_continuous(breaks = seq(0, 1, by = 0.2))+
  theme(axis.text = element_text(size = 5))

@@

```{r fig_sim, fig.height=4, fig.width=8}
ggarrange(fig_fgfpca, fig_adglmm,
          nrow = 1, common.legend = T)
```

@@

```{r}
## ISE container 
ise_mat <- ise_mat_ref <- array(NA, 
                                dim = c(length(window)-2, length(window)-2, M))
# dims: prediction window, max obs time, simulation iter

@@

```{r ise_fGFPCA}
## calculation
for(m in 1:M){
  this_df <- pred_list_all[[m]]
  ise_tb <- pred_list_all[[m]] %>%
    mutate(err1 = (pred0.2-eta_i)^2,
           err2 = (pred0.4-eta_i)^2,
           err3 = (pred0.6-eta_i)^2,
           err4 = (pred0.8-eta_i)^2) %>%
    select(id, t, starts_with("err")) %>% 
    mutate(window = cut(t, breaks = window, include.lowest = T)) %>% 
    group_by(window, id) %>% 
    summarise_at(vars(err1, err2, err3, err4), sum) %>% 
    group_by(window) %>% 
    summarize_at(vars(err1, err2, err3, err4), mean) %>%
    filter(window != "[0,0.2]") %>% 
    select(starts_with("err"))
  ise_mat[, ,m] <- as.matrix(ise_tb)

@@

mean_ise <- apply(ise_mat, c(1, 2), mean)

# mean_ise <- data.frame(mean_ise) %>%
#   mutate(Window = c("(0.2, 0.4]", "(0.4, 0.6]", "(0.6, 0.8]", "(0.8, 1.0]"),
#          .before = 1)
colnames(mean_ise) <- c("0.2", "0.4", "0.6", "0.8")
```

```{r ise_GLMMadaptive}
## calculation
for(m in 1:M){
  this_df <- pred_list_ref[[m]]
  ise_tb <- pred_list_ref[[m]] %>%
    mutate(err1 = (pred0.2-eta_i)^2,
           err2 = (pred0.4-eta_i)^2,
           err3 = (pred0.6-eta_i)^2,
           err4 = (pred0.8-eta_i)^2) %>%
    select(id, t, starts_with("err")) %>% 
    mutate(window = cut(t, breaks = window, include.lowest = T)) %>% 
    group_by(window, id) %>% 
    summarise_at(vars(err1, err2, err3, err4), sum) %>% 
    group_by(window) %>% 
    summarize_at(vars(err1, err2, err3, err4), mean) %>%
    filter(window != "[0,0.2]") %>% 
    select(starts_with("err"))
  ise_mat_ref[, ,m] <- as.matrix(ise_tb)

mean_ise_ref <- apply(ise_mat_ref, c(1, 2), mean)
# mean_ise_ref <- data.frame(mean_ise_ref) %>% 
#   mutate(Window = c("(0.2, 0.4]", "(0.4, 0.6]", "(0.6, 0.8]", "(0.8, 1.0]"),
#          .before = 1)
colnames(mean_ise_ref) <- c("0.2", "0.4", "0.6", "0.8")
```

```{r}
data.frame(Window = c("(0.2, 0.4]", "(0.4, 0.6]", "(0.6, 0.8]", "(0.8, 1.0]"),
           mean_ise, mean_ise_ref, check.names = F) %>%
  kable(digits = 3, caption = "Integrated squared error", booktabs=T) %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" "=1, "fGFPCA" = 4, "GLMMadaptive" = 4)) %>%
  add_header_above(c(" "= 1, "Maximum observation time" = 8))
```

@@

```{r}
## auc container 
auc_mat <- array(NA, dim = c(length(window)-2, length(window)-2, M))
auc_mat_ref <- array(NA, dim = c(length(window)-2, length(window)-2, M))
```

```{r auc_fGFPCA}
for(m in 1:M){
  this_df <- pred_list_all[[m]]
  auc_tb <- this_df %>% 
    mutate(window = cut(t, breaks = window, include.lowest = T)) %>% 
    select(Y, starts_with("pred"), window) %>%
    group_by(window) %>%
    summarise(auc1 = get_auc(Y, pred0.2),
              auc2 = get_auc(Y, pred0.4),
              auc3 = get_auc(Y, pred0.6),
              auc4 = get_auc(Y, pred0.8)) %>%
    filter(window != "[0,0.2]") %>% 
    select(starts_with("auc"))
  auc_mat[, ,m] <- as.matrix(auc_tb)

@@

mean_auc <- apply(auc_mat, c(1, 2), mean)
# mean_auc <- data.frame(mean_auc) %>% 
#   mutate(Window = c("(0.2, 0.4]", "(0.4, 0.6]", "(0.6, 0.8]", "(0.8, 1.0]"),
#          .before = 1)
colnames(mean_auc) <- c("0.2", "0.4", "0.6", "0.8")
```

```{r auc_GLMMadaptive}
for(m in 1:M){
  this_df <- pred_list_ref[[m]]
  auc_tb <- this_df %>% 
    mutate(window = cut(t, breaks = window, include.lowest = T)) %>% 
    select(Y, starts_with("pred"), window) %>%
    group_by(window) %>%
    summarise(auc1 = get_auc(Y, pred0.2),
              auc2 = get_auc(Y, pred0.4),
              auc3 = get_auc(Y, pred0.6),
              auc4 = get_auc(Y, pred0.8)) %>%
    filter(window != "[0,0.2]") %>% 
    select(starts_with("auc"))
  auc_mat_ref[, ,m] <- as.matrix(auc_tb)

@@

mean_auc_ref <- apply(auc_mat_ref, c(1, 2), mean)
# mean_auc <- data.frame(mean_auc) %>% 
#   mutate(Window = c("(0.2, 0.4]", "(0.4, 0.6]", "(0.6, 0.8]", "(0.8, 1.0]"),
#          .before = 1)
colnames(mean_auc_ref) <- c("0.2", "0.4", "0.6", "0.8")
```

```{r}
data.frame(Window = c("(0.2, 0.4]", "(0.4, 0.6]", "(0.6, 0.8]", "(0.8, 1.0]"),
           mean_auc, mean_auc_ref, check.names = F) %>%
  kable(digits = 3, caption = "Area under the ROC curve", booktabs = T) %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" "=1, "fGFPCA" = 4, "GLMMadaptive" = 4)) %>%
  add_header_above(c(" "= 1, "Maximum observation time" = 8))
```

```{r time}
data.frame(
  "Method" = c("fGFPCA", "GLMMadaptive"),
  "Fit" = c(mean(fit_time)/60, mean(fit_time_ref)),
           "Prediction"= c(mean(pred_time), mean(pred_time_ref)/60)) %>%
  kable(digits = 3, caption = "Computation time (minutes)", booktabs = T) %>%
  kable_styling(full_width = F) 
```

I think we could say that while the total time spend on model fitting + prediction are similar between two methods, fGFPCA achieved much better flexibility and much better predictive performance of prediction under every scenario.

@@

Here we would like to fit fGFPCA and GLMMadaptive on a dataset with smaller sample size and/or smaller measurement density. For the GLMMadaptive model, we would set it up with spline basis functions so that its flexibility is comparable with fGFPCA model, such as:

I have used 100 subjects for training and testing, and repeated 100 times. When fitting GLMMadpative, I reduce the number of measurements to 1/10 by taking one every 10 observations. The prediction is on the original grid.

```{r load_subset_sim_output}
load(here("Data/SubSimOutput_GLMMadaptive.RData"))
load(here("Data/SubSimOutput_fGFPCA.RData"))
```

@@

```{r fig_pred_subset}
rand_id2 <- sample(unique(pred_subset_fGFPCA[[1]]$id), 4)

fig_fgfpca_subset <- pred_subset_fGFPCA[[1]] %>%
  filter(id %in% rand_id2) %>%
  mutate_at(vars(eta_i, starts_with("pred")), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.6, col = "0.6"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), linetype="dashed", na.rm=T)+
  facet_wrap(~id)+
  labs(title = "fGFPCA", col = "Maximum observation time")+
  scale_x_continuous(breaks = seq(0, 1, by = 0.2))+
  theme(axis.text = element_text(size = 5))

fig_adglmm_subset <- pred_subset_adglmm[[1]] %>% 
  filter(id %in% rand_id2) %>%
  mutate_at(vars(eta_i, starts_with("pred")), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.6, col = "0.6"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), linetype="dashed", na.rm=T)+
  facet_wrap(~id)+
  labs(title = "GLMMadaptive", col = "Maximum observation time")+
  scale_x_continuous(breaks = seq(0, 1, by = 0.2))+
  theme(axis.text = element_text(size = 5))
```

```{r fit_sim_small, fig.height=4, fig.width=8}
ggarrange(fig_fgfpca_subset, fig_adglmm_subset,
          nrow = 1, common.legend = T)
```

>>>>

## prediction window 
window <- seq(0, 1, by = 0.2)
M <- 500
```

@@

@@

# Reference method

## GLMMadaptive

- For large datasets, we can fit a model with random intercept and slope for time. It is doable on 500 datasets, but obviously too simple for the data generation scheme. We would expect it to perform terribly.

- For small datasets, we would like to fit fGFPCA and GLMMadaptive on a dataset with smaller sample size and/or smaller measurement density. For the GLMMadaptive model, we would set it up with spline basis functions so that its flexibility is comparable with fGFPCA model, such as:

## Generalized functional-on-scalar regression

We would like to use a second reference method for predictive performance comparision. In this model, predictions on a specific interval are all made using the last few observations in the observed window. For example, if we are given observations on [0, 0.2] (j = 1...200), we may use L observations taken right before t=0.2 ($t_j: j=200, ... 200-(L-1)$) as time-fixed covariate to predict any future time.

Under both frameworks, for each dataset we need to fit four models (similar to the GLMMadaptive method):

But prediction performance will be reported by equal-length time window.

@@

@@

```{r load_sim1}
load(here("Data/SimOutput_fGFPCA.RData"))
load(here("Data/SimOutput_GLMMadaptive.RData"))
load(here("Data/SimOutput_GFOSR_L1.RData"))
pred_list_gfosr_l1 <- pred_list_gfofr
t_vec_gfosr_l1 <- t_vec_gfofr
rm(pred_list_gfofr, t_vec_gfofr)
load(here("Data/SimOutput_GFOSR_L5.RData"))
pred_list_gfosr_l5 <- pred_list_gfofr
t_vec_gfosr_l5 <- t_vec_gfofr
rm(pred_list_gfofr, t_vec_gfofr)

rand_id <- sample(501:600, size = 2)
```

```{r}
# range(pred_list_all[[1]]$pred0.2, na.rm = T)
fig_fgfpca <- pred_list_all[[1]] %>%
  filter(id %in% rand_id) %>%
  mutate_at(vars(eta_i, pred0.2, pred0.4, pred0.6, pred0.8), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.6, col = "0.6"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), linetype="dashed", na.rm=T)+
  facet_wrap(~id, nrow = 1)+
  labs(title = "fGFPCA", col = "Maximum observation time", x = "Time", y="")+
  scale_x_continuous(breaks = seq(0, 1, by = 0.2))+
  theme(axis.text = element_text(size = 5), 
        title = element_text(size = 8))

fig_adglmm <- pred_list_ref[[1]] %>%
  filter(id %in% rand_id) %>%
  mutate_at(vars(eta_i, pred0.2, pred0.4, pred0.6, pred0.8), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.6, col = "0.6"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), linetype="dashed", na.rm=T)+
  facet_wrap(~id, nrow = 1)+
  labs(title = "GLMMadaptive", col = "Maximum observation time", 
       x = "Time", y="")+
  scale_x_continuous(breaks = seq(0, 1, by = 0.2))+
  theme(axis.text = element_text(size = 5), 
        title = element_text(size = 8))

fig_gfosr_l1 <- pred_list_gfosr_l1[[1]] %>%
  filter(id %in% rand_id) %>%
  mutate_at(vars(eta_i, pred_w1, pred_w2, pred_w3, pred_w4), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  geom_line(aes(x=t, y=pred_w1, col = "0.2"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred_w2, col = "0.4"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred_w3, col = "0.6"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred_w4, col = "0.8"), linetype="dashed", na.rm=T)+
  facet_wrap(~id, nrow = 1)+
  labs(title = "GFOSR (L=1)", col = "Maximum observation time",
       x="Time", y="")+
  scale_x_continuous(breaks = seq(0, 1, by = 0.2))+
  theme(axis.text = element_text(size = 5), 
        title = element_text(size = 8))

fig_gfosr_l5 <- pred_list_gfosr_l5[[1]] %>%
  filter(id %in% rand_id) %>%
  mutate_at(vars(eta_i, pred_w1, pred_w2, pred_w3, pred_w4), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  geom_line(aes(x=t, y=pred_w1, col = "0.2"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred_w2, col = "0.4"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred_w3, col = "0.6"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred_w4, col = "0.8"), linetype="dashed", na.rm=T)+
  facet_wrap(~id, nrow = 1)+
  labs(title = "GFOSR (L=5)", col = "Maximum observation time",
       x="Time", y="")+
  scale_x_continuous(breaks = seq(0, 1, by = 0.2))+
  theme(axis.text = element_text(size = 5), 
        title = element_text(size = 8))

@@

```{r fig_sim, fig.height=8, fig.width=5}
ggarrange(fig_fgfpca, fig_adglmm, fig_gfosr_l1, fig_gfosr_l5,
          ncol = 1, common.legend = T)
```

@@

```{r}
## ISE container 
ise_fgfpca <- ise_adglmm <- ise_gfosr_l1 <- ise_gfosr_l5 <- 
  array(NA, dim = c(length(window)-2, length(window)-2, M))
# dims: prediction window, max obs time, simulation iter

@@

```{r ise_fGFPCA}
## calculation
for(m in 1:M){
  # this_df <- pred_list_all[[m]]
  ise_tb <- pred_list_all[[m]] %>%
    mutate(err1 = (pred0.2-eta_i)^2,
           err2 = (pred0.4-eta_i)^2,
           err3 = (pred0.6-eta_i)^2,
           err4 = (pred0.8-eta_i)^2) %>%
    select(id, t, starts_with("err")) %>% 
    mutate(window = cut(t, breaks = window, include.lowest = T)) %>% 
    group_by(window, id) %>% 
    summarise_at(vars(err1, err2, err3, err4), sum) %>% 
    group_by(window) %>% 
    summarize_at(vars(err1, err2, err3, err4), mean) %>%
    filter(window != "[0,0.2]") %>% 
    select(starts_with("err"))
  ise_fgfpca[, ,m] <- as.matrix(ise_tb)

@@

mean_ise_fgfpca <- apply(ise_fgfpca, c(1, 2), mean)

# mean_ise <- data.frame(mean_ise) %>%
#   mutate(Window = c("(0.2, 0.4]", "(0.4, 0.6]", "(0.6, 0.8]", "(0.8, 1.0]"),
#          .before = 1)
colnames(mean_ise_fgfpca) <- c("0.2", "0.4", "0.6", "0.8")
```

```{r ise_adglmm}
## calculation
for(m in 1:M){
  # this_df <- pred_list_gfosr_l1[[m]]
  ise_tb <- pred_list_ref[[m]] %>%
    mutate(err1 = (pred0.2-eta_i)^2,
           err2 = (pred0.4-eta_i)^2,
           err3 = (pred0.6-eta_i)^2,
           err4 = (pred0.8-eta_i)^2) %>%
    select(id, t, starts_with("err")) %>% 
    mutate(window = cut(t, breaks = window, include.lowest = T)) %>% 
    group_by(window, id) %>% 
    summarise_at(vars(err1, err2, err3, err4), sum) %>% 
    group_by(window) %>% 
    summarize_at(vars(err1, err2, err3, err4), mean) %>%
    filter(window != "[0,0.2]") %>% 
    select(starts_with("err"))
  ise_adglmm[, ,m] <- as.matrix(ise_tb)
}

mean_ise_adglmm <- apply(ise_adglmm, c(1, 2), mean)
colnames(mean_ise_adglmm) <- c("0.2", "0.4", "0.6", "0.8")
```

```{r ise_gfosr_l1}
## calculation
for(m in 1:M){
  # this_df <- pred_list_ref[[m]]
  ise_tb <- pred_list_gfosr_l1[[m]] %>%
    mutate(err1 = (pred_w1-eta_i)^2,
           err2 = (pred_w2-eta_i)^2,
           err3 = (pred_w3-eta_i)^2,
           err4 = (pred_w4-eta_i)^2) %>%
    select(id, t, starts_with("err"), window) %>% 
    # mutate(window = factor(window, levels = 1:5, 
    #                        labels = c("[0,0.2]", "(0.2,0.4]", "(0.4,0.6]",))) %>% 
    group_by(window, id) %>% 
    summarise_at(vars(err1, err2, err3, err4), sum) %>% 
    group_by(window) %>% 
    summarize_at(vars(err1, err2, err3, err4), mean) %>%
    filter(window != 1) %>% 
    select(starts_with("err"))
  ise_gfosr_l1[, ,m] <- as.matrix(ise_tb)
}

mean_ise_gfosr_l1 <- apply(ise_gfosr_l1, c(1, 2), mean)
colnames(mean_ise_gfosr_l1) <- c("0.2", "0.4", "0.6", "0.8")
```

```{r ise_gfosr_l5}
## calculation
for(m in 1:M){
  # this_df <- pred_list_ref[[m]]
  ise_tb <- pred_list_gfosr_l5[[m]] %>%
    mutate(err1 = (pred_w1-eta_i)^2,
           err2 = (pred_w2-eta_i)^2,
           err3 = (pred_w3-eta_i)^2,
           err4 = (pred_w4-eta_i)^2) %>%
    select(id, t, starts_with("err"), window) %>% 
    # mutate(window = factor(window, levels = 1:5, 
    #                        labels = c("[0,0.2]", "(0.2,0.4]", "(0.4,0.6]",))) %>% 
    group_by(window, id) %>% 
    summarise_at(vars(err1, err2, err3, err4), sum) %>% 
    group_by(window) %>% 
    summarize_at(vars(err1, err2, err3, err4), mean) %>%
    filter(window != 1) %>% 
    select(starts_with("err"))
  ise_gfosr_l5[, ,m] <- as.matrix(ise_tb)
}

mean_ise_gfosr_l5 <- apply(ise_gfosr_l5, c(1, 2), mean)
colnames(mean_ise_gfosr_l5) <- c("0.2", "0.4", "0.6", "0.8")
```

```{r}
rbind(mean_ise_fgfpca, mean_ise_adglmm, 
           mean_ise_gfosr_l1, mean_ise_gfosr_l5) %>%
  as_tibble() %>%
  mutate(Window = rep(c("(0.2, 0.4]", "(0.4, 0.6]", 
                        "(0.6, 0.8]", "(0.8, 1.0]"), 4),
         .before=1) %>%
  kable(digits = 3, booktabs=T,
        table.attr="style=\"color:black;\"") %>%
  kable_styling(full_width = F) %>% 
  group_rows(index = c("fGFPCA" = 4, "GLMMadaptive" = 4,
                     "GFOSR (L=1)" = 4, "GFOSR (L=5)" = 4)) %>%
  add_header_above(c(" "= 1, "Maximum observation time" = 4))
```

@@

```{r}
## auc container 
auc_fgfpca <- auc_adglmm <- auc_gfosr_l1 <- auc_gfosr_l5 <- 
  array(NA, dim = c(length(window)-2, length(window)-2, M))
```

```{r auc_fGFPCA}
for(m in 1:M){
  this_df <- pred_list_all[[m]]
  auc_tb <- this_df %>% 
    mutate(window = cut(t, breaks = window, include.lowest = T)) %>% 
    select(Y, starts_with("pred"), window) %>%
    group_by(window) %>%
    summarise(auc1 = get_auc(Y, pred0.2),
              auc2 = get_auc(Y, pred0.4),
              auc3 = get_auc(Y, pred0.6),
              auc4 = get_auc(Y, pred0.8)) %>%
    filter(window != "[0,0.2]") %>% 
    select(starts_with("auc"))
  auc_fgfpca[, ,m] <- as.matrix(auc_tb)
}

mean_auc_fgfpca <- apply(auc_fgfpca, c(1, 2), mean)
# mean_auc <- data.frame(mean_auc) %>% 
#   mutate(Window = c("(0.2, 0.4]", "(0.4, 0.6]", "(0.6, 0.8]", "(0.8, 1.0]"),
#          .before = 1)
colnames(mean_auc_fgfpca) <- c("0.2", "0.4", "0.6", "0.8")
```

```{r auc_GLMMadaptive}
for(m in 1:M){
  this_df <- pred_list_ref[[m]]
  auc_tb <- this_df %>% 
    mutate(window = cut(t, breaks = window, include.lowest = T)) %>% 
    select(Y, starts_with("pred"), window) %>%
    group_by(window) %>%
    summarise(auc1 = get_auc(Y, pred0.2),
              auc2 = get_auc(Y, pred0.4),
              auc3 = get_auc(Y, pred0.6),
              auc4 = get_auc(Y, pred0.8)) %>%
    filter(window != "[0,0.2]") %>% 
    select(starts_with("auc"))
  auc_adglmm[, ,m] <- as.matrix(auc_tb)

@@

mean_auc_adglmm <- apply(auc_adglmm, c(1, 2), mean)
colnames(mean_auc_adglmm) <- c("0.2", "0.4", "0.6", "0.8")
```
```{r auc_gfosr_l1}
for(m in 1:M){
  this_df <- pred_list_gfosr_l1[[m]]
  auc_tb <- this_df %>% 
    select(Y, starts_with("pred"), window) %>%
    group_by(window) %>%
    summarise(auc1 = get_auc(Y, pred_w1),
              auc2 = get_auc(Y, pred_w2),
              auc3 = get_auc(Y, pred_w3),
              auc4 = get_auc(Y, pred_w4)) %>%
    filter(window != 1) %>% 
    select(starts_with("auc"))
  auc_gfosr_l1[, ,m] <- as.matrix(auc_tb)

@@

mean_auc_gfosr_l1 <- apply(auc_gfosr_l1, c(1, 2), mean)
colnames(mean_auc_gfosr_l1) <- c("0.2", "0.4", "0.6", "0.8")
```

```{r auc_gfosr_l5}
for(m in 1:M){
  this_df <- pred_list_gfosr_l5[[m]]
  auc_tb <- this_df %>% 
    select(Y, starts_with("pred"), window) %>%
    group_by(window) %>%
    summarise(auc1 = get_auc(Y, pred_w1),
              auc2 = get_auc(Y, pred_w2),
              auc3 = get_auc(Y, pred_w3),
              auc4 = get_auc(Y, pred_w4)) %>%
    filter(window != 1) %>% 
    select(starts_with("auc"))
  auc_gfosr_l5[, ,m] <- as.matrix(auc_tb)

mean_auc_gfosr_l5 <- apply(auc_gfosr_l5, c(1, 2), mean)
colnames(mean_auc_gfosr_l5) <- c("0.2", "0.4", "0.6", "0.8")
```

```{r}
rbind(mean_auc_fgfpca, mean_auc_adglmm, 
           mean_auc_gfosr_l1, mean_auc_gfosr_l5) %>%
  as_tibble() %>%
  mutate(Window = rep(c("(0.2, 0.4]", "(0.4, 0.6]", 
                        "(0.6, 0.8]", "(0.8, 1.0]"), 4),
         .before=1) %>%
  kable(digits = 3, booktabs=T,
        table.attr="style=\"color:black;\"") %>%
  kable_styling(full_width = F) %>% 
  group_rows(index = c("fGFPCA" = 4, "GLMMadaptive" = 4,
                     "GFOSR (L=1)" = 4, "GFOSR (L=5)" = 4)) %>%
  add_header_above(c(" "= 1, "Maximum observation time" = 4))
```

## Time (minutes)

I chose not to report fitting and prediction time separately because it is too difficult to tell them apart for the GFOSR models.

```{r}
data.frame(
  "Method" = c("fGFPCA", "GLMMadaptive", "GFOSR (L=1)", "GFOSR (L=5)"),
  "Time" = c(mean(fit_time/60+pred_time),
             mean(fit_time_ref+pred_time_ref/60),
             mean(t_vec_gfosr_l1/60), 
             mean(t_vec_gfosr_l5/60))) %>%
  kable(digits = 3, booktabs = T, table.attr = "style = \"color:black;\"") %>%
  kable_styling(full_width = F) 
```

@@

I have used 100 subjects for training and testing, and repeated 500 times. When fitting GLMMadpative, I reduce the number of measurements in the training dataset to 1/10 by taking one every 10 observations. The prediction is on the original grid.

```{r load_sim2}
load(here("Data/SubSimOutput_GLMMadaptive.RData"))

load(here("Data/SubSimOutput_fGFPCA.RData"))
pred_subset_adglmm <- pred_subset_adglmm[!num_probs]
fit_time_subset_adglmm <- fit_time_subset_adglmm[!num_probs]
pred_time_subset_adglmm <- pred_time_subset_adglmm[!num_probs]

load(here("Data/SubSimOutput_GFOSR_L1.RData"))
pred_subset_gfosr_l1 <- pred_list_gfofr_subset
t_subset_gfosr_l1 <- t_vec_gfofr_subset
rm(pred_list_gfofr_subset, t_vec_gfofr_subset)

load(here("Data/SubSimOutput_GFOSR_L5.RData"))
pred_subset_gfosr_l5 <- pred_list_gfofr_subset
t_subset_gfosr_l5 <- t_vec_gfofr_subset
rm(pred_list_gfofr_subset, t_vec_gfofr_subset)

rand_id <- sample(101:200, size = 2)
```

@@

```{r fig_pred_subset}
fig_fgfpca2 <- pred_subset_fGFPCA[[1]] %>%
  filter(id %in% rand_id) %>%
  mutate_at(vars(eta_i, starts_with("pred")), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.6, col = "0.6"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), linetype="dashed", na.rm=T)+
  facet_wrap(~id)+
  labs(title = "fGFPCA", col = "Maximum observation time",
       x = "Time", y="")+
  scale_x_continuous(breaks = seq(0, 1, by = 0.2))+
  theme(axis.text = element_text(size = 5), 
        title = element_text(size = 8))

fig_adglmm2 <- pred_subset_adglmm[[1]] %>% 
  filter(id %in% rand_id) %>%
  mutate_at(vars(eta_i, starts_with("pred")), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.6, col = "0.6"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), linetype="dashed", na.rm=T)+
  facet_wrap(~id)+
  labs(title = "GLMMadaptive", col = "Maximum observation time",
       x = "Time", y = "")+
  scale_x_continuous(breaks = seq(0, 1, by = 0.2))+
  theme(axis.text = element_text(size = 5), 
        title = element_text(size = 8))

fig_gfosr_l1_2 <- pred_subset_gfosr_l1[[1]] %>% 
  filter(id %in% rand_id) %>%
  mutate_at(vars(eta_i, starts_with("pred")), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  geom_line(aes(x=t, y=pred_w1, col = "0.2"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred_w2, col = "0.4"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred_w3, col = "0.6"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred_w4, col = "0.8"), linetype="dashed", na.rm=T)+
  facet_wrap(~id)+
  labs(title = "GFOSR (L=1)", col = "Maximum observation time",
       x = "Time", y = "")+
  scale_x_continuous(breaks = seq(0, 1, by = 0.2))+
  theme(axis.text = element_text(size = 5), 
        title = element_text(size = 8))
